The Weakly Nonlinear Magnetorotational Instability in a local geometry

### 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
rearranged eqns

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{The weakly nonlinear magnetorotational instability}
\shortauthors{Clark \& Oishi}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\beq
\label{eqset2}
\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \Co B_0\partial_z B_{y} \, + \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
\eeq

\beq
\label{eqset3}
\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A
\eeq

\beq
\label{eqset4}
\partial_t B_{y} \, = \, B_0 \partial_z u_{y} \, + \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right) \, \\
+ \, \frac{1}{\reym} \nabla^2 B_{y}  \, - \, q \Omega_0 \partial_z A,
\eeq
\twocolumngrid

@@

Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}. To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI). We do so by substituting \edit1{$B = B_0\left(1 + \epsilon^2\right)$}. The degree of departure from the marginal state is measured by the small parameter $\epsilon$. An $\mathcal{O}\left(\epsilon^2\right)$ \edit1{strengthening} of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly. \edit1{We note that this definition of $\epsilon$ is opposite in sign to nearly all previous works \citep[e.g.][]{Umurhan:2007dz,Umurhan:2007hs}. Because in the ideal limit, the MRI can be tuned into instability by setting $B_0$ to its critical value and then \emph{decreasing} its value, it is natural to consider $\epsilon^2$ as a weakening of the background field. However, as we show in figure~\ref{fig:growth_rate_vs_B0}, for the dissipative case with $\eta, \nu \ne 0$, when all other parameters are critical, decreasing $B_0$ leads to stability, while increasing it pushes the system into instability. Figure~\ref{fig:growth_rate_vs_B0} is symmetric about $B_0 = 0$, as it must be, since the MRI is insensitive to the sign of the background field. There are several places at which the derivative of $\gamma$ appears discontinuous; this is not physical but rather reflects the fact that we define $\gamma$ as the growth rate of the most unstable mode. That is, it is the maximum real part of the eigenvalues of the linearized system (e.g. equation~\ref{eq:unperturbed_matrix_equations} with $\mathbf{N} = 0$). Because there are four wave families in rotating incompressible MHD, each modified differently by changing $B_0$, when the growth rates of the individual modes cross, there appear piecewise continuous solutions. We highlight one such point in the inset in figure~\ref{fig:growth_rate_vs_B0}, where the MRI mode becomes more stable than another mode which is always stable. Since all of these piecewise discontinuities are below $\gamma=0$, they do not affect the analysis here.}

\begin{figure}[[h!]]
  \centering
  \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
  \caption{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$. Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability. The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}
  \label{fig:growth_rate_vs_B0}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{The weakly nonlinear magnetorotational instability in a local geometry}
\shortauthors{Clark \& Oishi}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\beq
\label{eqset2}
\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, - \, \Co B_0\partial_z B_{y} \, - \, \frac{1}{\reye} \nabla^2 u_{y} \, = \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) 
\eeq

\beq
\label{eqset3}
\partial_t A \, - \, B_0 \partial_z \Psi \, - \, \frac{1}{\reym} \nabla^2 A \, = \, J\left(A, \Psi\right) 
\eeq

\beq
\label{eqset4}
\partial_t B_{y} \, + \, q \Omega_0 \partial_z A \, - \, B_0 \partial_z u_{y} \, - \, \frac{1}{\reym} \nabla^2 B_{y} \, = \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right)  ,
\eeq
\twocolumngrid

@@

Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}. To excite the weakly nonlinear MRI, we tune the background magnetic field away from stability. %(recall that stronger vertical fields stabilize the MRI). 
We do so by substituting \edit1{$B = B_0\left(1 + \epsilon^2\right)$}. The degree of departure from the marginal state is measured by the small parameter $\epsilon$. An $\mathcal{O}\left(\epsilon^2\right)$ \edit1{strengthening} of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly. \edit1{We note that this definition of $\epsilon$ is opposite in sign to nearly all previous works \citep[e.g.][]{Umurhan:2007dz,Umurhan:2007hs}. Because in the ideal limit, the MRI can be tuned into instability by setting $B_0$ to its critical value and then \emph{decreasing} its value, it is natural to consider $\epsilon^2$ as a weakening of the background field. However, as we show in figure~\ref{fig:growth_rate_vs_B0}, for the dissipative case with $\eta, \nu \ne 0$, when all other parameters are critical, decreasing $B_0$ leads to stability, while increasing it pushes the system into instability. Figure~\ref{fig:growth_rate_vs_B0} is symmetric about $B_0 = 0$, as it must be, since the MRI is insensitive to the sign of the background field. There are several places at which the derivative of $\gamma$ appears discontinuous; this is not physical but rather reflects the fact that we define $\gamma$ as the growth rate of the most unstable mode. That is, it is the maximum real part of the eigenvalues of the linearized system (e.g. equation~\ref{eq:unperturbed_matrix_equations} with $\mathbf{N} = 0$). Because there are four wave families in rotating incompressible MHD, each modified differently by changing $B_0$, when the growth rates of the individual modes cross, there appear piecewise continuous solutions. We highlight one such point in the inset in figure~\ref{fig:growth_rate_vs_B0}, where the MRI mode becomes more stable than another mode which is always stable. Since all of these piecewise discontinuities are below $\gamma=0$, they do not affect the analysis here.}

\begin{figure}[[h!]]
  \centering
  \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
  \caption{\edit1{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$. Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability. The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}}
  \label{fig:growth_rate_vs_B0}
\end{figure}
